Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-09-23 packages/spec-haskell/yellowpaper/Paper.tex
WIP Payment Execution Environment Models

\subsubsection{Using Shared Context}

But the discrete nature of the model does not provide the necessary element for us to add the desired properties to the
payment system. Here we propose a modification that involves the usage of \textit{shared context (ctx for short)}:

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{../assets/money-distribution-with-ctx.png}
    \caption{Schematic representation of money distribution with shared context}
    \label{fig:md-with-ctx}
\end{figure}

Note that in this model, shared context can be updated independently while value functions of the same money
distribution can produce different monetary values.

@@

\begin{itemize}
\item If time is included the shared context, then monetary value of each monetary unit can be continuous over time.
\item A subset of monetary units can also have their value functions depending on the same information in shared
    context. This could enable payment primitives that involve many parties.
\end{itemize}

\subsubsection{Haskell Definitions}

@@

Payment primitives are functions that takes a money distribution, shared context and some arbitrary arguments as inputs
and an updated money distribution and shared context as outputs:

\begin{minted}[escapeinside=||,mathescape=true]{haskell}
    somePaymentPrimitive :: MoneyDistribution |$\otimes$| SharedContext
                         -> Args
                         -> MoneyDistribution |$\otimes$| SharedContext
\end{minted}

Loosely speaking, it is considered a primitive, if it can not be broken down into other existing primitives which result
in the same money distribution.

The best known primitive is instant transfer of monetary value between one monetary unit to another. Shared context
enables more primitives to be defined, which will be introduced in later chapters.

\subsection{Payment System Models}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Discrete Payments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Continuous Payments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\subsubsection{Using Context}

But the discrete nature of the model does not provide the necessary element for us to add the desired properties to the
payment system. Here we propose a modification that involves the usage of \textit{context ($\gamma$)}:

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{../assets/money-distribution-with-ctx.png}
    \caption{Schematic representation of money distribution with context}
    \label{fig:md-with-ctx}
\end{figure}

Note that in this model, context can be updated independently while value functions of the same money distribution can
produce different monetary values.

@@

\begin{itemize}
\item If time (t) is included the context, then monetary value of each monetary unit can be continuous over time. Time
    is part of the physical reality, hence not changeable by the actors in the payment system.
\item Any subset of monetary units can also have their value functions depending on the same information in
    context. This could enable payment primitives that involve many parties. This set of information in context is
    referred to as \textit{ctx :: SharedContext}, they can be changed over time by the actors in the payment system.
\end{itemize}

For the purpose of this paper, the model of context is: $\gamma = (t :: Timestamp) \oplus (ctx :: SharedContext)$.

\subsubsection{Haskell Definition Of Money Distribution}

@@

Payment primitives are functions that takes a money distribution, shared context, some payment instruction (args) and
timestamp as inputs and incremental updates\footnote{Specifically being monoidal, that is in short a set that has
associative binary operation and an identity element. See \url{https://ncatlab.org/nlab/show/monoid}.} of money
distribution and shared context as outputs:

\begin{minted}[escapeinside=||,mathescape=true]{haskell}
    somePaymentPrimitive :: ( Timestamp t
                            , MoneyDistribution md
                            , SharedContext ctx
                            , PaymentInstruction args
                            , Monoid md, Monoid ctx
                            )
                         => md |$\otimes$| ctx -> args -> t -> md |$\otimes$| ctx
\end{minted}

Loosely speaking, it is considered a primitive, if it can not be broken down into other existing primitives which result
in the same money distribution; additionally primitives should be the only constructs in a payment system that can
update money distribution.

Updates are \textit{monoidal}, so that they can be incremental and their parallel executions can be modeled.

The best known primitive is instant transfer of monetary value between one monetary unit to another. The introduction of
context enables more primitives to be defined, and this will be discussed in later chapters.

\subsection{Payment Execution Environment Models}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{FinancialContract.lhs.tex}
